### 2021-09-14 article/article.tex
update

DeepCore is a dense sub-array whithin the standard IceCube array, which allows us to observe neutrino interactions down to a few GeV~\cite{DC2021}.
Due to its lower energy threshold compared with IceCube, DeepCore has previously been used to study NSI effects in this more discernible range\cite{DC2021,deepcoreNSI}.

PINGU is a proposed 26 string in-fill array to IceCube, consisting of optical modules of similar nature~\cite{PINGUletter}. While porposed to mainly shed light on 
neutrino mass ordering, PINGU will be able to probe NSI parameters aswell.  
\subsection{IceCube}\label{ch:ICmethod}

We obtain the data from the IC-86 sterile data release~\cite{IC2020}, which contain muon track events collected over 8 years. 
In the data, the reconstructed energy $\Ereco$ is logarithmically binned between \SI{500}{\GeV} to \SI{9976}{\GeV}, totalling 13 bins (index $i$).
The reconstructed cosine-zenith $\zreco$ is linearly binned between $-1$ and $0$ in 20 bins (index $j$).

@@

Now we turn to the effect of $\emt$. Muon events are the most abundant, and it suffices to study $P_{\mu\mu}$ in this way to explain the $\emt$ features. 
First, let $P^\pm$ denote $P_{\mu\mu}(\epsilon^\pm_{\mu\tau})$. 
We see see that $\emt^+$ generates a slightly higher $P_{\mu\mu}$ for energies around \SI{20}{\GeV} (blue area),
while $\emt^-$ produces higher $P_{\mu\mu}$ for almost all other combinations of ${\Etrue,\ztrue}$. As we see in Fig.~\ref{fig:flux_ratio}, this muon survival abundance 
is indeed preserved at flux level. So the flux for $\emt^+$ is higher than the flux for $\emt^-$. Is this still true at event level, i.e. after reconstruction? 
As we see in Fig.~\ref{fig:emt_events}, the binned PINGU event counts display strong differences for many bins, which will give a high statistics on both sides of $\emt=0$, slightly favoring $\emt^-$. 
DeepCore on the other hand, has fewer bins where the event count for $\emt^-$ surpasses the event count for $\emt^+$, 
giving weaker statistics for the negative side. Thus, we conclude that we will see a $\emt$ \si{GeV}-asymmetry stemming from the lower statistics 
for negative $\emt$ at probability level, which is then propagated through the flux and finally affects the reconstruction. %TODO: fill more conclusion?

>>>>

DeepCore is a dense sub-array within the standard IceCube array, which allows us to observe neutrino interactions down to a few GeV~\cite{DC2021}.
Due to its lower energy threshold compared with IceCube, DeepCore has previously been used to study NSI effects in this more discernible range\cite{DC2021,deepcoreNSI}.

PINGU is a proposed 26 string in-fill array to IceCube, consisting of optical modules of similar nature~\cite{PINGUletter}. While proposed to mainly shed light on 
neutrino mass ordering, PINGU will be able to probe NSI parameters as well.  
\subsection{IceCube}\label{ch:ICmethod}

We obtain the data from the IC-86 sterile data release~\cite{IC2020}, which contains muon track events collected over 8 years. 
In the data, the reconstructed energy $\Ereco$ is logarithmically binned between \SI{500}{\GeV} to \SI{9976}{\GeV}, totalling 13 bins (index $i$).
The reconstructed cosine-zenith $\zreco$ is linearly binned between $-1$ and $0$ in 20 bins (index $j$).

@@

Now we turn to the effect of $\emt$. Muon events are the most abundant, and it suffices to study $P_{\mu\mu}$ in this way to explain the $\emt$ features. 
First, let $P^\pm$ denote $P_{\mu\mu}(\epsilon^\pm_{\mu\tau})$. 
We see see that $\emt^+$ generates a slightly higher $P_{\mu\mu}$ for energies around \SI{20}{\GeV} (blue area),
while $\emt^-$ produces higher $P_{\mu\mu}$ for almost all other combinations of ${\Etrue,\ztrue}$. As we see in Fig.~\ref{fig:flux_ratio}, this muon survival abundance 
is indeed preserved at flux level. So the flux for $\emt^+$ is higher than the flux for $\emt^-$. Is this still true at event level, i.e. after reconstruction? 
As we see in Fig.~\ref{fig:emt_events}, the binned PINGU event counts display strong differences for many bins, which will give a high statistics on both sides of $\emt=0$, slightly favoring $\emt^-$. 
DeepCore on the other hand has fewer bins where the event count for $\emt^-$ surpasses the event count for $\emt^+$, 
giving weaker statistics for the negative side. Thus, we conclude that we will see a $\emt$ \si{GeV}-asymmetry stemming from the lower statistics 
for negative $\emt$ at probability level, which is then propagated through the flux and finally affects the reconstruction. %TODO: fill more conclusion?

### 2021-09-14 thesis/tex/DCtheory.tex
update

The oscillation parameters used on our DeepCore simulations are from the
best-fit in the global analysis in~\cite{nufit}: $\theta_{12} = \SI{33.44}{\degree},\, \theta_{13} = \SI{8.57}{\degree},\, \Delta m^2_{21} =  \SI{7.42}{\electronvolt^2}$, and we 
marginalize over $\dm$ and $\theta_{23}$.

>>>>

The oscillation parameters used on our DeepCore simulations are from the
best-fit in the global analysis in~\cite{nufit}: $\theta_{12} = \SI{33.44}{\degree},\, \theta_{13} = \SI{8.57}{\degree},\, \Delta m^2_{21} =  \SI{7.42}{\electronvolt^2}$, and we 
marginalize over $\dm[31]$ and $\theta_{23}$.

### 2021-09-14 thesis/tex/ICintro.tex
update

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{graphicx, slashed, siunitx, subcaption}
% \usepackage[utf8]{inputenc}
% \usepackage{xr-hyper}
% \input{commands}
% \begin{document}
\section{Neutrino detection}
We always observe neutrinos indirectly through their associated charged lepton. 
Regardless of the type of interaction (charged current via the $W$ boson, or neutral current
via the $Z$), a charged lepton exits with altered properties. The lepton is then detected, and the properties of the neutrino involved in the 
interaction is then deduced. This deduction will be imperfect, and this introduces complexities that we will handle in Ch.~\ref{ch:ICmethod}.

In this work, we only study the detectors handeled by the IceCube collaboration. They are of Cherenkov type, which means that they detect 
the secondary charged lepton by its emitted Cherenkov light, produced from its travel through the Antarctic ice. 
If the charged leptons interact heavily with the ice, they will travel a short distance and emit a localized flash of 
Cherenkov light. This event is referred to as a cascade. The neutral current interactions involves quarks, which recoils and procudes
showers of hadrons. Also, charged current $\ne$ interactions also produce cascades. A cascade event 
is shown in Fig.~\ref{fig:events_cascade}.
If the charged leptons don't interact as much in the ice, they penetrate a larger part of it, emitting light and tertiary particles
as they go. This event is referred to as a track, and are often due to muon charged current interactions. A track event 
is shown in Fig.~\ref{fig:events_track}.

@@

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than other.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilites of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
results from IceCube and DeepCore. The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

\begin{figure}[t]\label{fig:flux}
    \includegraphics[width=0.7\textwidth]{figures/flux.pdf}
    \caption{Interpolated South Pole atmospheric $\nu_\mu$ flux with data from~\cite{hondaArticle}, displaying a near-neglible 
    zenith angle dependance, but a strong energy dependence. The energy dependence can be parametrized to be approximated by $E^{-2.7}$.}
\end{figure}

\subsection{Atmospheric Neutrino Flux}
Atmospheric neutrinos originates from cosmic rays composed of high-energy protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector.
The detector's experience an uncorrelated systematic error, which comes from this \emph{muon background}, 
i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. For the DeepCore analysis, we will have to consider this background when calculating the events.
For IceCube events, we will scan a higher energy range where the muon background can be neglected. For the PINGU events, the IceCube detector is 
expected to be able to act as a veto for this background. Thus, the error introduced from the muon background is expected by the collaboration 
to be neglible~\cite{PINGUletter}.
Moreover, the atmospheric flux itself is often associated with a large error.
In this work, we will use a flux normalization error of 25\%, and a zenith slope error of 5\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}.
The flux data is binned in $\cos(\theta_z)$. The fluxes are averaged over azimuthal direction and over solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities on the form 
$P_{\tau \beta}$, since we have no starting atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of approximating all four necessary fluxes for a given true energy and true zenith.
The result of $\phi_\mu$ is shown in Fig.~\ref{fig:flux}. We see that the flux is almost entirely zenith independent. This is due to the 
very small cross-section of the neutrinos. For particles interacting through the electromagnetic interaction, we would expect to see a 
large difference between the flux of particles traversing through the whole diameter of the Earth, compared to particles that only penetrate a 
small rim of the crust. Here, however, the flux at the South Pole is almost the same! This does not mean that the oscillation probability will be 
equal. First, the difference in the density profiles introduces wildly different matter effects. 
Secondly, the zenith angle determines the baseline. Completely down-going neutrinos, with $\cos(\theta_z) = 0$ will be produced at an average distance of \SI{20}{\km}
overhead, travel down to the icy sheet of Antarctica, and then penetrate the ice down \SI{2}{\km} to the DOMs. 
The energy dependence of the flux is approximately $E^{-2.7}$, so our expected event count at higher energies will suffer from this strong decline in flux.

So now we have

@@

\subsection{Event Reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. 
This means that, given the parameters recorded by the detector, what are their "true" values?
We are interested in two variables: the energy and the direction. Each event is tagged with a probable energy and zenith angle, 
called the recostructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. 
So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true energy and zenith angle, $\Etrue$ and $\ztrue$. 
The reconstructed parameters are what we are using to analyze the data (because this is what the detector actually sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).

>>>>

% \documentclass{thesis}
% \usepackage[margin=0.5in]{geometry}
% \usepackage{graphicx, slashed, siunitx, subcaption}
% \usepackage[utf8]{inputenc}
% \usepackage{xr-hyper}
% \input{commands}
% \begin{document}
\section{Neutrino detection}
We always observe neutrinos indirectly through their associated charged lepton. 
Regardless of the type of interaction (charged current via the $W$ boson, or neutral current
via the $Z$), a charged lepton exits with altered properties. The lepton is then detected, and the properties of the neutrino involved in the 
interaction are then deduced. This deduction will be imperfect, and this introduces complexities that we will handle in Ch.~\ref{ch:ICmethod}.

In this work, we only study the detectors handled by the IceCube collaboration. They are of Cherenkov type, which means that they detect 
the secondary charged lepton by its emitted Cherenkov light, produced from its travel through the Antarctic ice. 
If the charged leptons interact heavily with the ice, they will travel a short distance and emit a localized flash of 
Cherenkov light. This event is referred to as a cascade. The neutral current interactions involve quarks, which recoils and produces
showers of hadrons. Also, charged current $\ne$ interactions also produce cascades. A cascade event 
is shown in Fig.~\ref{fig:events_cascade}.
If the charged leptons don't interact as much in the ice, they penetrate a larger part of it, emitting light and tertiary particles
as they go. This event is referred to as a track and is often due to charged current muon interactions. A track event 
is shown in Fig.~\ref{fig:events_track}.

@@

The strings and DOMs are not spaced evenly, making some parts of the detector more sensitive to certain energy ranges than others.
8 strings packed more tightly than the other 78, making that part of the detector sensitive to neutrino energies down to single-digit \si{\GeV}. Due to 
this part being situated deep within the ice, it is referred to DeepCore. DeepCore will be treated as a separate and independent detector from the rest, which
retains the name IceCube. A view of the current setup can be seen in Fig.~\ref{fig:array}. In this work, we consider DeepCore data between \SIrange{5.6}{56}{\GeV} and IceCube data in the range \SIrange{0.5}{10}{\TeV}.
\begin{figure}\label{fig:array}
    \centering
    \includegraphics[width=0.5\textwidth]{figures/icecube2.png}
    \caption{View of the full IceCube array, with the Eiffel tower for scale.}
\end{figure}

In 2017, the PINGU Letter of Intent was published~\cite{PINGUletter}. The `Precision IceCube Next Generation Upgrade' is an upgrade that will 
supplement DeepCore, i.e. boosing the capabilities of neutrino detection at the \si{\GeV} scale. As the PINGU upgrade is not yet financed nor built, we are
not able to use any data from it. However, the collaboration has released preliminary simulations which we will use to see how the upgrade might improve
results from IceCube and DeepCore. The PINGU simulations have the same structure as the DeepCore data, so our analysis referring to DeepCore will
also apply to PINGU except where noted. However, we treat the PINGU detector as independent of the other experiments.

\subsection{Atmospheric Neutrino Flux}
Atmospheric neutrinos originate from cosmic rays composed of high-energy protons interacting with nuclei in the atmosphere.
These interactions ultimately produces pions, which decay as 
\begin{align}\label{eq:pion}
    &\pi^+ \to \mu^+ + \nm\,, \quad \pi^- \to \mu^- + \anm \nonumber \\
    &\pi^+ \to e^+ + \ne\,,\,\, \quad \pi^- \to e^- + \ane\,.
\end{align}
In the muonic decay channel, muons are emitted which will be detected by the IceCube detector.
The detector's experience an uncorrelated systematic error, which comes from this \emph{muon background}, 
i.e. events misclassified as muons from 
$\nm$ interactions rather than from pion decay. For the DeepCore analysis, we will have to consider this background when calculating the events.
For IceCube events, we will scan a higher energy range where the muon background can be neglected. For the PINGU events, the IceCube detector is 
expected to be able to act as a veto for this background. Thus, the error introduced from the muon background is expected by the collaboration 
to be neglible~\cite{PINGUletter}.
Moreover, the atmospheric flux itself is often associated with a large error.
In this work, we will use a flux normalization error of 25\%, and a zenith slope error of 5\%~\cite{hondapaper}.

The flux is provided in \cite{hondaData,hondaArticle}, and a selection is shown in Table~\ref{table:flux}.
The flux data is binned in $\cos(\theta_z)$. The fluxes are averaged over the azimuthal direction and solar minimum/maximum. 
The units of the fluxes are given as \si{\per\GeV \per\metre\squared \per\second \per\steradian} and are omitted
from the table for clarity. 
We note that the fluxes for $\nt$ and $\nu_{\bar{\tau}}$ are missing. Kaons can decay into neutral pions, which in turn 
can produce $\nt$, but this branching ratio is extremely small. Thus, we never have to use probabilities in the form 
$P_{\tau \beta}$, since we have no starting atmospheric $\nu_\tau$ flux. 
Interpolating the data yields makes us capable of approximating all four necessary fluxes for a given true energy and true zenith.
The result of $\phi_\mu$ is shown in Fig.~\ref{fig:flux}. We see that the flux is almost entirely zenith independent. This is due to the 
very small cross-section of the neutrinos. For particles interacting through the electromagnetic interaction, we would expect to see a 
large difference between the flux of particles traversing through the whole diameter of the Earth, compared to particles that only penetrate a 
small rim of the crust. Here, however, the flux at the South Pole is almost the same! This does not mean that the oscillation probability will be 
equal. First, the difference in the density profiles introduces wildly different matter effects. 
Secondly, the zenith angle determines the baseline. Completely down-going neutrinos, with $\cos(\theta_z) = 0$ will be produced at an average distance of \SI{15}{\km}
overhead, travel down to the icy sheet of Antarctica, and then penetrate the ice down \SI{2}{\km} to the DOMs~\cite{hondapaper}. 
The energy dependence of the flux is approximately $E^{-2.7}$, so our expected event count at higher energies will suffer from this strong decline in flux.

@@

\begin{figure}[t]\label{fig:flux}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/flux.pdf}
    \caption{Interpolated South Pole atmospheric $\nu_\mu$ flux with data from~\cite{hondaArticle}, displaying a near-neglible 
    zenith angle dependance, but a strong energy dependence. The energy dependence can be parametrized to be approximated by $E^{-2.7}$.}
\end{figure}

We propagate the atmospheric neutrino flux $\phi_\alpha^\text{atm}(E,\theta_z)$ through the Earth.
The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux for flavor $\beta$ at detector level as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$.

\subsection{Event Reconstruction}
After an event has occurred, the IceCube algorithms process the data coming from the detector to \emph{reconstruct} the event. 
This means that, given the parameters recorded by the detector, what are their `true' values?
We are interested in two variables: energy and direction. Each event is tagged with a probable energy and zenith angle, 
called the reconstructed parameters $\Ereco$ and $\zreco$, which are the parameters according to the DOMs.
The collaboration then uses numerous sophisticated methods to backtrack the reconstructed parameters to the true parameters. 
So a charged lepton hits the DOMs, and we ultimately end up with the associated neutrino's true energy and zenith angle, $\Etrue$ and $\ztrue$. 
The reconstructed parameters are what we are using to analyze the data (because this is what the detector sees), while the true parameters are used in the 
determination of that neutrino's `actual' flux and cross-section (because this is what Nature sees).
